Technical Interview

### 2017-09-04 tex/paper/paper.tex
Near final version of the paper

\newpage
\tableofcontents* % the asterisk means that the contents itself isn't put into the ToC
%\clearpage %or \cleardoublepage
\phantomsection
\listoffigures
%\listoftables
\newpage

@@

%-------------------------------------------------------------------------------
\section{The Problem}
\label{sec:problem}
We interpret the question as the following the mathematical problem. 
We define a $X$ as a discrete random variable. 
As input, we have a $k$ values, $x_i$, with their associated probabilities, $p_i$.

@@

We then use a bounded uniform distribution, $U\left(0,1\right)$,
to generate random numbers between $0$ and $1$.
We picked two pseudo random numbers generator : Mersenne Twister\footnote{
The Mersenne twister is a pseudorandom number generator developed by 
Makoto Matsumoto and Takuji Nishimura that is based on a matrix linear 
recurrence over a finite binary field. 
It has a period of $2^{63}-1$.
The source for the C code can be found 
\href{http://www.math.sci.hiroshima-u.ac.jp/~m-mat/MT/emt64.html}{here}.
} (This implementation claims to be faster than Ecuyer's one.) and Ecuyer\footnote{
This is the default random number generator in haskell one in haskell.
 This implementation uses the Portable Combined Generator of L'Ecuyer
 for 32-bit computers, transliterated in Haskell by Lennart Augustsson. 
 It has a period of roughly $2.30584e18$.
 } in order to compare and pick the best one.

@@

	In this module we have the implementation of the discrete 
	probability function\footnote{
	We can take as input incomplete probabilities, i.e. that
	don't sum up to $1$. We therefore assign the remaining
	probability to the event we don't select any element.
	Haskell provides the \emph{Maybe a} type to modelise this.
	}, the discrete cumulative function, 
	the inverse discrete cumulative function and the mean and
	standard deviation functions.

@@

\subsection{User Interface}
The UI is a command line input allowing the user to run each feature.
After building the project, with the following command:
\begin{verbatim}
cabal build
\end{verbatim}
The binary is then located in \emph{./dist/build/MAN-AHL/MAN-AHL}.
We can run the weighted or the uniform engine
 as described in \autoref{fig:cli}, with the following command:
\begin{verbatim}
time ./dist/build/MAN-AHL/MAN-AHL -run=Weighted -nSims=1000000
        -pillars=''
\end{verbatim}
 The binary has the option \emph{-help}, stating all the default values.
\begin{figure}[h!]
\begin{center}
\includegraphics[width=1\textwidth]{img/cli.png}
\caption{Command Line Input}
\label{fig:cli}
\end{center}
\end{figure}

@@

The aim was to keep the software as simple as possible while targetting a
good design and performance. 
The right balance is always difficult to achieve as unfortunately no 
software is ever perfect.

The code is documented while we believe that a simple and well organised
code needs less documentation. The complexities are presented in the key
area of the code. The code is presented in \autoref{apx:code}.

@@

We subsequently define the ADTs to represent all the curves we need
for our weighted probability engine :
\begin{itemize}
	\item The \emph{PDF} ADT modelising the probability mass function,
	The key are the $x_i$ and the values the $p_i$, we presented in
	\autoref{sec:problem}.
	We merge the pillars with similar $x_i$, discard the pillars with zero
	$p_i$ and ensure that $\sum_i p_i \leq 1$.
	If needed we a a pillar representing non selected element 
	as \emph{Nothing}. This way we ensure that the probabilities
	adds up to 1.
	\item The \emph{CDF} ADT represents the discrete cumulative
	curve, presented in \autoref{fig:cdf2}.
	\item The \emph{InvCDF} ADT represents the inverse CDF curve,
	as explained in \autoref{eq:invcdf}
	\item The \emph{Distribution a} ADT models a distribution curve
	of the random number generated.
\end{itemize}
Again, we refer to curve as to piecewise curves or histograms.

@@

In a way, we lose static typing here in order to merge all the types of 
request and to dispatch the query to the appropriate engine. And we merge
back the result into the \emph{Result} to provide a single function achieving
everything we need.

@@

The following Test Suites, i.e. lists of tests, compose our test plan split by modules.

@@

\begin{description}
	\item [Mean and Standard Deviation]
		Given a large number of random number generated for
		both Ecuyer and Mersenne, i.e. $100000$.
		We ensure that the mean and the standard deviation 
		of the distributions are still the same. Run 10 times.
		We ensure that the uniform random number engines keep
		the same properties.
	\item [Bounds]
		Test ensuring that the uniform pseudo random generators
		generate probabilities with the $[0, 1]$ bound.
		Run $10$ times for $100000$ random numbers.
	\item [Uniform]
		Ensure that the ditributions generated are uniform,
		by making sure that we match a flat distribution at 0.001
		precision.
		Run 10 times for $100000$ random numbers.
	\item [Performance]
		Ensure that the pseudo random numbers keep at
		least the same performance.
		Run 10 times for $100000$ random numbers.
\end{description}

\subsection{Weighted Engine Testing} 
These are the unit tests for the weighted pseudo random engine.
All tests are run using the two available uniform pseudo random
generators. 
These are like the integration tests as they use most of the library.
\begin{description}
	\item [Distribution check]
		We check that we recover the PDF distribution from
		a sample of 1000000 random numbers at $0.1\%$ precision.
		We run this for 10 pseudo random PDF.
	\item [Engine Fail]
		Ensure that the construction of the engine 
		handles all the expected failure.
	\item [Performance]
		Ensure that the engine keep at least the same performance.
		Run 10 times for $100000$ random  numbers	
\end{description}

@@

%-------------------------------------------------------------------------------
\section{Qualitative Analysis}
We realise that we maybe be beyond the scope of the initial
question, which was to implement a \emph{nextNum} function.
However how to test it efficiently ? It is a very hard question.

It is an important part of Software Engineering and 
Quantitative Development to understand the 
meaning of what we are trying to achieve. 
It allows better modelisation, algoritms and testing. 
In this section, we will look at the number generated and try
to find the best parameters for this software.

\subsection{Metrics}
To assess the quality of a random number engine, we use
simple statistics. It would have been nice to have time to 
study further statistical and mathematical papers related 
to the random numbers generation and probabilities.
We could have created with more advanced metrics.

There are two keys elements which drive the results :
the input discrete probability function with $k$ pillars and
the output discrete probability function deducted from a
simulation of $n$ random numbers.

@@

We also have two random numbers generator, 
Mersenne and Ecuyer, available. We compare them thoughout.

As our weighted probability engine depends on
the uniform probability engine. 
We test first the uniform engine.

The metrics we look at are:
\begin{itemize}
	\item the difference between the input and output pdf,
	\item the standard deviation of these differences,
	\item the mean of these difference, hopefully close to zero.
\end{itemize}

@@

\begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{img/uni-m-1.png}
        \caption{A gull}
        \label{fig:gull}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{img/uni-m-1.png}
        \caption{A tiger}
        \label{fig:tiger}
    \end{subfigure}

    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{img/uni-m-1.png}
        \caption{A gull}
        \label{fig:gull}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{img/uni-m-1.png}
        \caption{A tiger}
        \label{fig:tiger}
    \end{subfigure}

    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{img/uni-m-1.png}
        \caption{A gull}
        \label{fig:gull}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{img/uni-m-1.png}
        \caption{A tiger}
        \label{fig:tiger}
    \end{subfigure}
    \caption{Pictures of animals}\label{fig:animals}
\end{figure}

@@

@@

\subsection{Complexity Analysis}

\subsection{Uniform Engine Performance}

@@

%-------------------------------------------------------------------------------
\section{Conclusion}

%-------------------------------------------------------------------------------
\section{Possible Extensions}
Here is a list of possible extension we could add.
\begin{description}
	\item [Analytics]
	Further abstract the PDF, CDF, InvCDF to handle all kind of keys.
	\item [Code Design]
	We wonder if there was a way to avoid the \emph{MultiParamTypeClasses}
	extension and to achieve a better design. 
	Unfortunately time didn't allow further explorations.
	\item [Testing]
	We could run the tests on different environments to ensure the portability of 
	the code.
	\item [UI] 
	A nice extension would be to add a Graphic User Interface,
	using the library 
	\href{https://hackage.haskell.org/package/threepenny-gui}{threepenny-gui}
	to create a portable Web application.

@@

%-------------------------------------------------------------------------------
\newpage
\appendix
\addcontentsline{toc}{chapter}{Appendices}

\chapter{Code} \label{apx:code}
We provide below the code. It is screenshots from 
\href{https://vim.sourceforge.io/}{Vim} with the 
\href{https://github.com/begriffs/haskell-vim-now}{haskell-vim-now}.
extension.

%-------------------------------------------------------------------------------
\newpage
\chapter{Installation} \label{apx:installation}

Here are the steps to install the App.
\begin{enumerate}
	\item Download Haskell via this \href{https://www.haskell.org/platform/}{link}.
	\item Install following the instructions given by the installer.
	\item Use cabal to download the require packages :
\begin{itemize}
	\item mersenne-random-pure64
	\item countainers
	\item random
	\item mtl
	\item QuickCheck
	\item deepseq
\end{itemize}
\end{enumerate}
\end{document}

>>>>

\newpage
\tableofcontents* % the asterisk means that the contents itself isn't put into the ToC
%\clearpage %or \cleardoublepage
\phantomsection
\listoffigures
\listoftables
\vfill
\newpage

@@

%-------------------------------------------------------------------------------
\section{The Problem}
\label{sec:problem}
We interpret the question as the following the mathematical problem. 
We define $X$ as a discrete random variable. 
As input, we have a $k$ values, $x_i$, with their associated probabilities, $p_i$.

@@

We then use a bounded uniform distribution, $U\left(0,1\right)$,
to generate random numbers between $0$ and $1$.
We picked two pseudo random numbers generator : Mersenne Twister\footnote{
The Mersenne twister is a pseudorandom number generator developed by 
Makoto Matsumoto and Takuji Nishimura that is based on a matrix linear 
recurrence over a finite binary field. 
It has a period of $2^{63}-1$.
The source for the C code can be found 
\href{http://www.math.sci.hiroshima-u.ac.jp/~m-mat/MT/emt64.html}{here}.
} (This implementation claims to be faster than Ecuyer's one.) and Ecuyer\footnote{
This is the default random number generator in haskell.
 This implementation uses the Portable Combined Generator of L'Ecuyer
 for 32-bit computers, transliterated in Haskell by Lennart Augustsson. 
 It has a period of roughly $2.30584e18$.
 } in order to compare and pick the best one.

@@

	In this module we have the implementation of the discrete 
	probability function\footnote{
	We can take as input incomplete probabilities, i.e. that
	don't sum up to $1$. We therefore assign the remaining
	probability to the event we don't select any element.
	Haskell provides the \emph{Maybe a} type to modelise this.
	}, the discrete cumulative function, 
	the inverse discrete cumulative function, the mean and
	standard deviation functions.

@@

\subsection{User Interface}
The UI is a command line input allowing the user to run each feature.
After building the project, with the following command:
\begin{verbatim}
cabal build
\end{verbatim}
The binary is then located in \emph{./dist/build/MAN-AHL/MAN-AHL}.
We can run the weighted or the uniform engine
 as described in \autoref{fig:cli}, with the following command:
\begin{verbatim}
time ./dist/build/MAN-AHL/MAN-AHL -run=Weighted -nSims=1000000
        -pillars='[(1, 0.5), (2, 0.5)]'
\end{verbatim}
 The binary has the option \emph{-help}, stating all the default values.
\begin{figure}[h!]
\begin{center}
\includegraphics[width=1\textwidth]{img/cli.png}
\caption{Command Line Input}
\label{fig:cli}
\end{center}
\end{figure}

@@

The aim was to keep the software as simple as possible while targeting a
good design and performance. 
The right balance is always difficult to achieve as unfortunately no 
software is ever perfect.

The code is documented while we believe that a simple and well organised
code needs less documentation. The complexities are presented in the key
area of the code.

@@

We subsequently define the ADTs to represent all the curves we need
for our weighted probability engine :
\begin{itemize}
	\item The \emph{PDF} ADT modelising the probability mass function,
	The key are the $x_i$ and the values the $p_i$, we presented in
	\autoref{sec:problem}.
	We merge the pillars with similar $x_i$, discard the pillars with zero
	$p_i$ and ensure that $\sum_i p_i \leq 1$.
	If needed we add a pillar representing non selected element 
	as \emph{Nothing}. This way we ensure that the probabilities
	adds up to 1.
	\item The \emph{CDF} ADT represents the discrete cumulative
	curve, presented in \autoref{fig:cdf2}.
	\item The \emph{InvCDF} ADT represents the inverse CDF curve,
	as explained in \autoref{eq:invcdf}
	\item The \emph{Distribution a} ADT models a distribution curve
	of the random number generated.
\end{itemize}
Again, we refer to curve as to piecewise curves or histograms.

@@

In a way, we lose static typing here in order to merge all the types of 
requests and to dispatch the query to the appropriate engine. And we merge
back the result into the \emph{Result} to provide a single function achieving
everything we need.

@@

The following Test Suites, i.e. lists of tests, compose our test plan split by modules.
There is more on the testing in \autoref{sec:quality}. We expand on what is a correct 
metric for this problem.

@@

\begin{description}
	\item [Mean and Standard Deviation]
		Given a large number of random number generated for
		both Ecuyer and Mersenne, i.e. $100000$.
		We ensure that the mean and the standard deviation 
		of the distributions are still the same. Run 10 times.
		We ensure that the uniform random number engines keep
		the same properties.
	\item [Bounds]
		Test ensuring that the uniform pseudo random generators
		generate probabilities with the $[0, 1]$ bound.
		Run $10$ times for $1000000$ random numbers.
	\item [Uniform]
		Ensure that the ditributions generated are uniform,
		by making sure that we match a flat distribution at 0.001
		precision on the standard deviation of the difference
		between the PDFs.
		Run 10 times for $100000$ random numbers.
	\item [Performance]
		Ensure that the pseudo random numbers keep at
		least the same performance.
		Run 10 times for $100000$ random numbers.
\end{description}

\subsection{Weighted Engine Testing} 
These are the unit tests for the weighted pseudo random engine.
All tests are run using the two available uniform pseudo random
generators. 
These are like the integration tests as they use most of the library.
\begin{description}
	\item [Distribution check]
		We check that we recover the PDF distribution from
		a sample of 1000000 random numbers at $0.1\%$ precision
		on the standard deviation of the differences with the 
		generated PDF.
		We run this for 10 pseudo random PDF.
	\item [Engine Fail]
		Ensure that the construction of the engine 
		handles all the expected failure.
	\item [Performance]
		Ensure that the engine keep at least the same performance.
		Run 10 times for $100000$ random  numbers	
\end{description}

@@

%-------------------------------------------------------------------------------
\section{Qualitative Analysis}
\label{sec:quality}
We realise that we maybe be beyond the scope of the initial
question, which was to implement a \emph{nextNum} function.
However how to test it efficiently ? It is a very hard question.
This led us to look at the distribution of the results to ensure
the quality of the numbers generated.

In this section, we will look at the number generated and try
to find the best parameters for this software.

\subsection{Metrics}
To assess the quality of a random number engine, we use
simple statistics.

There are two keys elements that drive the results :
the input discrete probability function with $k$ pillars and
the output discrete probability function deducted from a
simulation of $n$ random numbers.

@@

We also have two random numbers generators, 
Mersenne and Ecuyer, available. We compare them thoughout.

As our weighted probability engine depends on
the uniform probability engine, we test first the uniform engine.

The metrics we look at are:
\begin{itemize}
	\item the difference between the input and output pdf,
	\item the standard deviation of these differences,
	\item the mean of these differences, hopefully close to zero,
	\item and the highest and lowest differences.
\end{itemize}

@@

\autoref{tab:uni.test} presents several runs for various number of random
numbers for each random number generators, which can be use in
the bounded uniform probability engine.
% Table generated by Excel2LaTeX from sheet 'Sheet5'
\begin{table}[htbp!]
  \centering \tiny
    \begin{tabular}{r|rrrrrr}
    \multicolumn{1}{r}{} & \multicolumn{2}{c}{1000000} & \multicolumn{2}{c}{100000} & \multicolumn{2}{c}{10000} \\
    \multicolumn{1}{r}{} & Mersenne & Ecuyer & Mersenne & Ecuyer & Mersenne & Ecuyer \\
\cmidrule{2-7}    Mean  & 5.55E-19 & 4.16E-19 & 4.86E-19 & 2.78E-19 & 5.55E-19 & 4.16E-19 \\
    StdDev  & 1.08E-04 & 1.27E-04 & 4.42E-04 & 4.80E-04 & 1.47E-03 & 1.29E-03 \\
    Highest  & 1.89E-04 & 2.89E-04 & 9.90E-04 & 1.01E-03 & 3.80E-03 & 2.50E-03 \\
    Lowest  & -2.91E-04 & -2.93E-04 & -1.56E-03 & -1.25E-03 & -3.10E-03 & -2.60E-03 \\
    \end{tabular}%
  \caption{Uniform Distribution Test}
  \label{tab:uni.test}%
\end{table}%

The best standard deviation we achieve is of the order of $e-4$.
The mean is very close to zero with both generators and the results 
are very similar.
Hence setting $n$ to 1000000 for more accurate testing, 
i.e. of the order of $e-4$, is recommanded.

We tried various $k$, less than 100 to remain realistic. It has no impact on the
results as expected.

@@

Therefore we can't expect to achieve better accuracy with the
weighted probability engine as it relies on the bounded uniform engine.

We present in \autoref{tab:w.test}, the result of multiple runs with various $n$.
% Table generated by Excel2LaTeX from sheet 'Sheet5'
\begin{table}[htbp!]
  \centering \tiny
    \begin{tabular}{r|rrrrrr}
    \multicolumn{1}{r}{} & \multicolumn{2}{c}{1000000} & \multicolumn{2}{c}{100000} & \multicolumn{2}{c}{10000} \\
    \multicolumn{1}{r}{} & Mersenne & Ecuyer & Mersenne & Ecuyer & Mersenne & Ecuyer \\
\cmidrule{2-7}    Mean  & -3.73E-18 & -3.36E-18 & -3.45E-18 & -3.96E-18 & -3.38E-18 & -3.26E-18 \\
    StdDev  & 1.71E-04 & 1.49E-04 & 5.14E-04 & 4.61E-04 & 1.79E-03 & 1.91E-03 \\
    Highest  & 7.88E-04 & 2.70E-04 & 1.13E-03 & 8.00E-04 & 5.20E-03 & 5.50E-03 \\
    Lowest  & -2.99E-04 & -3.61E-04 & -1.34E-03 & -1.25E-03 & -5.10E-03 & -5.52E-03 \\
    \end{tabular}%
  \caption{Weighted Distribution Test}
  \label{tab:w.test}%
\end{table}%
The result are very close to the precision of the uniform engine. 
We therefore recommand the same setting as before for $n$ and $k$.

@@

\subsection{Complexity Analysis}
We commented some of the code with the complexities.
The computation of the weighted distribution for $n$ random number generated, 
i.e. the function \emph{allStats}, with a PDF of $k$ pillars, is :
\[
	O(n . log (k) . rng )
\]
where $rng$ is the complexity of the raw random number generator.

\subsection{Uniform Engine Performance}
\autoref{tab:uni.perf} presents the result of many runs with various $n$ and
the heap that was allocated during the run and the time taken.
\begin{table}[htbp!]
  \centering \tiny
    \begin{tabular}{r|rrrrrr}
    \multicolumn{1}{r}{} & \multicolumn{2}{c}{1000000} & \multicolumn{2}{c}{100000} & \multicolumn{2}{c}{10000} \\
    \multicolumn{1}{r}{} & Mersenne & Ecuyer & Mersenne & Ecuyer & Mersenne & Ecuyer \\
\cmidrule{2-7}    Heap  & 2,724,561,112 & 3,360,007,920 & 274,635,272 & 338,190,648 & 29,602,472 & 35,953,944 \\
    Time  & 3.576 & 3.707 & 0.347 & 0.373 & 0.041 & 0.041 \\
    \end{tabular}%
  \caption{Uniform Engine Performance}
  \label{tab:uni.perf}%
\end{table}%
Most of the heap allocation and the time is consumed by the underlying 
random number generator, the various engines adds only minor overhead, 
as shown on the profiling results, c.f. \autoref{fig:perf1} and \autoref{fig:perf2}.
We can see that the time and space are linear with respect to $n$. 
\begin{figure}[h!]
\begin{center}
\includegraphics[width=1\textwidth]{img/prof1.png}
\caption{Performance Profiling 1}
\label{fig:perf1}
\end{center}
\end{figure}
\begin{figure}[h!]
\begin{center}
\includegraphics[width=1\textwidth]{img/prof2.png}
\caption{Performance Profiling 2}
\label{fig:perf2}
\end{center}
\end{figure}
%\begin{figure}[h!]
%\begin{center}
%\includegraphics[width=1\textwidth]{img/prof3.png}
%\caption{Performance Profiling 3}
%\label{fig:perf3}
%\end{center}
%\end{figure}
We can conclude that the complexity of the inner random number generators
are $O(1)$ in time and $O(n)$ in space. Indeed the function \emph{allStats} for the
\emph{StatUniEngine} does not store the number generated, 
just the distribution of these.

We notice that Ecuyer take much more space, therefore our default setting
is set to Mersenne.

@@

\autoref{tab:w.perf} presents similar results for the \emph{StatWPEngine}, 
i.e. the statistic weighted probability engine.
\begin{table}[htbp!]
  \centering \tiny
    \begin{tabular}{r|rrrrrr}
    \multicolumn{1}{r}{} & \multicolumn{2}{c}{1000000} & \multicolumn{2}{c}{100000} & \multicolumn{2}{c}{10000} \\
    \multicolumn{1}{r}{} & Mersenne & Ecuyer & Mersenne & Ecuyer & Mersenne & Ecuyer \\
\cmidrule{2-7}    Heap  & 2,777,630,784 & 3,419,213,784 & 281,175,032 & 344,423,280 & 30,695,560 & 37,067,344 \\
    Time  & 3.612 & 3.81  & 0.357 & 0.379 & 0.038 & 0.043 \\
    \end{tabular}%
  \caption{Weighted Engine Performance}
  \label{tab:w.perf}%
\end{table}%
The results shows that there is little overhead for the weighted probabilities
engines. The results are very similar to \autoref{tab:uni.perf}.
This conforts us in the computation of our complexities.

%-------------------------------------------------------------------------------
\section{Possible Extensions}
Here is a list of possible extension we could add.
\begin{description}
	\item [Analytics]
	Further abstract the PDF, CDF, InvCDF to handle all kind of keys.
	\item [Testing]
	We could run the tests on different environments to ensure the portability of 
	the code.
	\item [UI] 
	A nice extension would be to add a Graphic User Interface,
	using the library 
	\href{https://hackage.haskell.org/package/threepenny-gui}{threepenny-gui}
	to create a portable Web application.

@@

\end{document}
